Four decades of water quality change in the upper San Francisco Estuary

### 2017-01-11 sftrends_manu.Rnw
mods and trends data files updated with md10
added md10 data to text, combiend d7 clam figs

What do nitrogen trends mean?  Have to interpret relative to trends in other variables.  A decrease in nitrogen or constant nitrogen does not mean nitrogen inputs have stayed the same, they might actually be increasing if nitrogen.  A change in chlorohpyll relative to change in nitrogen could be informative, and even moreso, a change in silica relative to change in chlorophyll suggests diatom biomass has changed.  However, there are mismatches in these trends that suggest other processes are at play, e.g., residence times and flow inputs, etc.  Trends in Suisun relative to trends in Delta provide an example, e.g., Suisun is decrease in chlorophyll, increase in silica, increase in nitrogen, delta is decrease in silica, increase/decrease in DIN (depending on time period/season), decrease in chlorophyll, what's going on? See Senn slide 14 (from burial?). The WRTDS model lets us at least address trends in the context of season, time, and flow.  This allows for more improved interpretion relative to observing raw data. Also explain more information by looking at ammonium, nitrative/nitrite, relative to DIN. What about other variables (light level as suspended particulate matter, temperature)?

@@

# combine all
trnds_seasyr <- rbind(trnds_nrm, trnds_fit, trnds_obs) %>% 
  select(Site_Code, resvar, cat, ztest, tau, perchg, ann, trndvar) %>% 
  filter(!trndvar %in% 'fits') %>% 
  mutate(
    Site_Code = factor(Site_Code, levels = c('D7', 'D6', 'D4', 'D28', 'D26', 'D19', 'P8', 'C10', 'C3')),
    ztest = ifelse(ztest < 0.05, '(p < 0.05)', '(ns)'),
    resvar = factor(resvar, levels = reslabs$shrt, labels = reslabs$expr), 
    trndvar = factor(trndvar, levels = c('res', 'norm'), labels = c('Obs.', 'Flow-norm.'))
  ) %>% 
  unite('trnd', trndvar, ztest, sep = ' ')

@@

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for aggregations by (a) year, (b) seasons from 1976-1995, and (c) seasons from 1996-2013.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Values less than -9\\% change were truncated for visual clarity of the bar plots, see \\cref{tab:trndsann,tab:trndsbef,tab:trndsaft} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 5.3, cache = T, out.width = '0.7\\textwidth'>>=

@@

# combine all
trnds_seasyr <- rbind(trnds_nrm, trnds_fit, trnds_obs) %>% 
  select(Site_Code, resvar, cat, perchg, ann, trndvar) %>% 
  spread(trndvar, perchg) %>% 
  mutate(Site_Code = factor(Site_Code, levels = c('D7', 'D6', 'D4', 'D28', 'D26', 'D19', 'P8', 'C10', 'C3')))

@@

# need to shift locations of some stations
statmeta[statmeta$Site_Code %in% 'D26', 'Latitude'] <- 38.12
statmeta[statmeta$Site_Code %in% 'D26', 'Longitude'] <--121.55
statmeta[statmeta$Site_Code %in% 'D7', 'Latitude'] <- 38.13
statmeta[statmeta$Site_Code %in% 'D7', 'Longitude'] <--122.015 
statmeta[statmeta$Site_Code %in% 'C10', 'Latitude'] <-37.68 
statmeta[statmeta$Site_Code %in% 'C10', 'Longitude'] <--121.28

@@

# bounds on limits
yrng <- c(-9, 9)

@@

      nms <- names(bars)[i]
      dat <- statmeta[statmeta$Site_Code %in% nms,]
      pall <- subview(pall, bars[[i]], x = dat$Longitude, y = dat$Latitude, subwid, 0.148) # change bar heights here

@@

# setup legend
rngs <- seq(-9, 7, by = 3)
leg <- ggplot(data.frame(x = seq_along(rngs), y = seq_along(rngs), z = rngs), aes(x, y, colour = z)) + 
  geom_point() + 
  scale_colour_gradientn(colours = c(rev(colneg), colpos)) + 
  guides(colour = guide_colourbar(title = expression('% chg yr' ^-1), barheight = 0.5, barwidth = 8)) + 
  theme(legend.position = 'top')
leg <- g_legend(leg)

@@

# dynaplots
p5 <- dynaplot(nomod, logspace = F, col_vec = dynacol_vec, month = c(1, 4, 7, 10), floscl = F) + 
  scale_x_continuous(flolab) + 
  scale_y_continuous(ylab1) +
  ptheme %+replace% 
  theme(
    legend.position = 'top', 
    legend.title = element_blank()
  ) +
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10))
pleg2 <- g_legend(p5)
p5 <- p5 + ptheme %+replace% theme(legend.position = 'none')
p6 <- dynaplot(nhmod, logspace = F, col_vec = dynacol_vec, month = c(1, 4, 7, 10), floscl = F) + 
  scale_x_continuous(flolab) +
  scale_y_continuous(ylab2) +
  ptheme %+replace% 
  theme(
    legend.position = 'none', 
    legend.title = element_blank()
  )

@@

<<clmchl, fig.height = 7, fig.width = 8, out.width = '\\textwidth', cache = T, fig.cap = 'Trends in clam abundance and \\ac{chla} concentration from 1976 to 2014 at station D7 in Suisun Bay.  Invasion by \\textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \\textit{Corbicula fluminea} was shown by changes in clam density (a, annual means).  A coincident decrease in \\ac{chla} concentration was also observed (c).  A weak but significant ($p < 0.001$) relationship between clam biomass and \\ac{chla} concentration is shown in subfigure (b).  Flow relationships with \\ac{chla} concentration have also changed over time (d, observations from June). Chlorophyll shows a slight positive then dominantly negative association with increasing flow (decreasing salinity) early in the time series, whereas the trend is reversed in recent years.'>>=

# clam and model data at D7
data(clams)
data(h3dat)
data(flow_dat)
chld7 <- filter(h3dat, Site_Code %in% 'D7' & resvar %in% 'chl')$data[[1]]

@@

# subfig labs
txt <- data.frame(
  x = c(as.numeric(as.Date('1979-10-01')), 2500, 25, 0.25),
  y =  c(10000, 14, 4, 3.3), 
  txt = c('(a)', '(c)', '(b)', '(d)')
)
size <- 6

##
# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    # axis.title.x = element_blank(),
    legend.position = 'top', 
    legend.title = element_blank()
    ) + 
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('Year', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  with(txt[1, ], annotate('text', x = xlims[1] + 365 * 2, y = y, label = txt, size = size))

@@

# scatter plot of concentration x biomass with model 
p2 <- ggplot(clmflo, aes(x = q, y = biomass)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + biomass) = 5.2 - 0.6log(1 + flow)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0,15)) +
  geom_text(data = txt[2, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size)

##
# chlorophyll by month
chldat <- filter(h3dat, resvar %in% 'chl' & Site_Code %in% 'D7') %>% 
  .$data %>% 
  .[[1]] %>% 
  select(year, month, res, bt_norm) %>% 
  mutate(res = exp(res)) %>% 
  rename(yr = year, mo = month, bt_res = res) %>% 
  group_by(yr, mo) %>% 
  dplyr::summarize(res = mean(bt_res, na.rm = TRUE), bt_norm = mean(bt_norm, na.rm = TRUE)) %>% 
  ungroup

@@

xlab <- expression(paste(italic(Potamocorbula), " biomass (g ", m^-2, ")"))
ylab <- expression(paste("Flow-normalized chlorophyll (", italic(mu), "g ", L^-1, ")"))

# scatter plot of concetration x biomass with model 
p3 <- ggplot(toplo, aes(x = biomass, y = bt_norm)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal'
    ) +
  scale_colour_manual(label = 'log(1 + chl) = 1.35 - 0.1log(1 + biomass)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0.8, 4)) +
  geom_text(data = txt[3, ], aes(x = x, y = y, label = txt, colour = NULL), show.legend = F, size = size)

##
# dynaplot for chl, june
cols <- RColorBrewer::brewer.pal(10, 'Set2')[1:2]
ylab <- attr(chld7, 'reslab')

p4 <- dynaplot(chld7, mo = 6, col_vec = cols, floscl = F) +
  theme_minimal() +
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"),
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal',
    strip.background = element_blank(),
    strip.text.x = element_blank()
    ) + 
  scale_x_continuous('ln-Salinity (psu)') + 
  scale_y_continuous(ylab, limits = c(0.4, 3.3)) +
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10)) + 
  annotate('text', x = txt[4, 'x'], y = txt[4, 'y'], label = txt[4, 'txt'], size = size)

# align widths of plots in first column, first two
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p2))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth

# align widths of plots in first column, first two
pC <- ggplot_gtable(ggplot_build(p3))
pD <- ggplot_gtable(ggplot_build(p4))
maxWidth = unit.pmax(pC$widths[2:3], pD$widths[2:3])

# Set the widths
pC$widths[2:3] <- maxWidth
pD$widths[2:3] <- maxWidth

# combine all
grid.arrange(
  arrangeGrob(pA, pB, ncol = 1, heights = c(1, 1)), 
  arrangeGrob(pC, pD, ncol = 1, heights = c(1.05, 1)), 
  ncol = 2, widths = c(1, 1)
  )
@

<<d7c10trnds, fig.cap = 'Flow-normalized trends of annual (left) and seasonal (right) variation in DIN, chl-{\\it\\footnotesize a}, and SiO$_2$ at station D7 (top) and C10 (bottom).  Covariation between nutrients, chl-{\\it\\footnotesize a}, and SiO$_2$ is observed at D7 but not C10, although an overall decrease in SiO$_2$ at C10 is shown.  Seasonal changes at D7 are most pronounced during the summer.', fig.width = 9, fig.height = 10, cache = T>>=
data(h3dat)

# separate each site
toplo1 <- filter(h3dat, Site_Code %in% 'D7')
toplo2 <- filter(h3dat, Site_Code %in% 'C10')

# y axis labels
ylab1 <- expression(paste("DIN (mg ", L^-1, ")"))
ylab2 <- expression(paste("Chl-", italic(a), " (", italic(mu), "g ", L^-1, ")"))
ylab3 <- expression(paste("Si", O[2], " (mg ", L^-1, ")"))

# generic theme for all plots
ptheme <- theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"),  
    axis.title.x = element_blank(), 
    axis.text.x = element_blank()
  )

# colors
cols <- RColorBrewer::brewer.pal(9, 'Set1')[1:3]
col_vec <- RColorBrewer::brewer.pal(9, 'Blues')[c(9, 8, 7)]

## first set of plots
d7din<- filter(toplo1, resvar == 'din') %>% 
  .$data %>% 
  .[[1]]
pdin1 <- prdnrmplot(d7din, col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  scale_y_continuous(ylab1) +
  theme(legend.position = 'top', legend.title = element_blank())
pleg1 <- g_legend(pdin1)
pdin1 <- pdin1 + theme(legend.position = 'none')

pdinsy1 <- seasyrplot(d7din, col_vec = cols, logspace = F, predicted = F) +
  ptheme + 
  theme(legend.position = 'top', axis.title.y = element_blank()) + 
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10))
pleg2 <- g_legend(pdinsy1)
pdinsy1 <- pdinsy1 + theme(legend.position = 'none')

d7chl<- filter(toplo1, resvar == 'chl') %>% 
  .$data %>% 
  .[[1]]
pchl1 <- prdnrmplot(d7chl, col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none') + 
  scale_y_continuous(ylab2)

pchlsy1 <- seasyrplot(d7chl, col_vec = cols, logspace = F, predicted = F) +
  ptheme + 
  theme(legend.position = 'none', axis.title.y = element_blank())

d7sio2 <- filter(toplo1, resvar == 'sio2') %>% 
  .$data %>% 
  .[[1]]
psio1 <- prdnrmplot(d7sio2, col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text()) +
  scale_x_date('Year') +
  scale_y_continuous(ylab3)

psiosy1 <- seasyrplot(d7sio2, col_vec = cols, logspace = F, predicted = F) + 
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text(), axis.title.y = element_blank()) +
  scale_x_date('Day of year', date_labels = "%m/%d")

# Set the widths
pA1 <- ggplot_gtable(ggplot_build(pdin1))
pB1 <- ggplot_gtable(ggplot_build(pchl1))
pC1 <- ggplot_gtable(ggplot_build(psio1))
maxWidth = unit.pmax(pA1$widths[2:3], pB1$widths[2:3], pC1$widths[2:3])
pA1$widths[2:3] <- maxWidth
pB1$widths[2:3] <- maxWidth
pC1$widths[2:3] <- maxWidth

# Set the widths for seasyrplots
pAsy1 <- ggplot_gtable(ggplot_build(pdinsy1))
pBsy1 <- ggplot_gtable(ggplot_build(pchlsy1))
pCsy1 <- ggplot_gtable(ggplot_build(psiosy1))
maxWidth = unit.pmax(pAsy1$widths[2:3], pBsy1$widths[2:3], pCsy1$widths[2:3])
pAsy1$widths[2:3] <- maxWidth
pBsy1$widths[2:3] <- maxWidth
pCsy1$widths[2:3] <- maxWidth

## second set of plots
c10din <- filter(toplo2, resvar == 'din') %>% 
  .$data %>% 
  .[[1]]
pdin2 <- prdnrmplot(c10din, col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  scale_y_continuous(ylab1) +
  theme(legend.position = 'none')

pdinsy2 <- seasyrplot(c10din, col_vec = cols, logspace = F, predicted = F) +
  ptheme + 
  theme(legend.position = 'none', axis.title.y = element_blank())

c10chl <- filter(toplo2, resvar == 'chl') %>% 
  .$data %>% 
  .[[1]]
pchl2 <- prdnrmplot(c10chl, col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none') + 
  scale_y_continuous(ylab2)

pchlsy2 <- seasyrplot(c10chl, col_vec = cols, logspace = F, predicted = F) +
  ptheme + 
  theme(legend.position = 'none', axis.title.y = element_blank())

c10sio2 <- filter(toplo2, resvar == 'sio2') %>% 
  .$data %>% 
  .[[1]]
psio2 <- prdnrmplot(c10sio2, col_vec = col_vec, logspace = F, min_mo = 10) +
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text()) +
  scale_x_date('Year') +
  scale_y_continuous(ylab3)

psiosy2 <- seasyrplot(c10sio2, col_vec = cols, logspace = F, predicted = F) + 
  ptheme + 
  theme(legend.position = 'none', axis.title.x = element_text(), axis.text.x = element_text(), axis.title.y = element_blank()) +
  scale_x_date('Day of year', date_labels = "%m/%d")

# Set the widths
pA2 <- ggplot_gtable(ggplot_build(pdin2))
pB2 <- ggplot_gtable(ggplot_build(pchl2))
pC2 <- ggplot_gtable(ggplot_build(psio2))
maxWidth = unit.pmax(pA2$widths[2:3], pB2$widths[2:3], pC2$widths[2:3])
pA2$widths[2:3] <- maxWidth
pB2$widths[2:3] <- maxWidth
pC2$widths[2:3] <- maxWidth

# Set the widths for seasyrplots
pAsy2 <- ggplot_gtable(ggplot_build(pdinsy2))
pBsy2 <- ggplot_gtable(ggplot_build(pchlsy2))
pCsy2 <- ggplot_gtable(ggplot_build(psiosy2))
maxWidth = unit.pmax(pAsy2$widths[2:3], pBsy2$widths[2:3], pCsy2$widths[2:3])
pAsy2$widths[2:3] <- maxWidth
pBsy2$widths[2:3] <- maxWidth
pCsy2$widths[2:3] <- maxWidth

grid.arrange(
  arrangeGrob(
    ncol = 1, heights = c(0.05, 1),
    arrangeGrob(pleg1, pleg2, ncol = 2),
    arrangeGrob(
      arrangeGrob(
        textGrob('(a) Station D7, Suisun Bay'), heights = c(0.1, 1), 
        arrangeGrob(
          arrangeGrob(pA1, pB1, pC1, ncol = 1, heights = c(0.85, 0.85, 1)),
          arrangeGrob(pAsy1, pBsy1, pCsy1, ncol = 1, heights = c(0.85, 0.85, 1)),
          ncol = 2
        )
      ),
      arrangeGrob(
        textGrob('(b) Station C10, Delta'), heights = c(0.1, 1),
        arrangeGrob(
          arrangeGrob(pA2, pB2, pC2, ncol = 1, heights = c(0.85, 0.85, 1)),
          arrangeGrob(pAsy2, pBsy2, pCsy2, ncol = 1, heights = c(0.85, 0.85, 1)),
          ncol = 2
        )
      ),
      ncol = 1 
    )
  )
)

@@

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Annual'),
  n.cgroup = c(2),
  rowname = stats,
  label = 'tab:trndsann',
  insert.bottom = foot.val
  )

@@

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Seasonal, 1976-1995'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsbef',
  insert.bottom = foot.val
  )

@@

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = c(9, 9, 9),
  cgroup = c('Seasonal, 1996-2013'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsaft',
  insert.bottom = foot.val
  )

@@

% trends in din, chl, sio2 at c10
<<c10chg>>=

c10dat <- filter(h3dat, Site_Code %in% 'C10')

dinraw <- filter(c10dat, resvar %in% 'din') %>% 
  .$data %>% 
  .[[1]]
chlraw <- filter(c10dat, resvar %in% 'chl') %>% 
  .$data %>% 
  .[[1]]
sioraw <- filter(c10dat, resvar %in% 'sio2') %>% 
  .$data %>% 
  .[[1]]

>>>>

What do nitrogen trends mean?  Have to interpret relative to trends in other variables.  A decrease in nitrogen or constant nitrogen does not mean nitrogen inputs have stayed the same, they might actually be increasing if nitrogen.  A change in chlorophyll relative to change in nitrogen could be informative, and even moreso, a change in silica relative to change in chlorophyll suggests diatom biomass has changed.  However, there are mismatches in these trends that suggest other processes are at play, e.g., residence times and flow inputs, etc.  Trends in Suisun relative to trends in Delta provide an example, e.g., Suisun is decrease in chlorophyll, increase in silica, increase in nitrogen, delta is decrease in silica, increase/decrease in DIN (depending on time period/season), decrease in chlorophyll, what's going on? See Senn slide 14 (from burial?). The WRTDS model lets us at least address trends in the context of season, time, and flow.  This allows for more improved interpretion relative to observing raw data. Also explain more information by looking at ammonium, nitrative/nitrite, relative to DIN. What about other variables (light level as suspended particulate matter, temperature)?

@@

# combine all
trnds_seasyr <- rbind(trnds_nrm, trnds_fit, trnds_obs) %>% 
  select(Site_Code, resvar, cat, ztest, tau, perchg, ann, trndvar) %>% 
  filter(!trndvar %in% 'fits') %>% 
  mutate(
    Site_Code = factor(Site_Code, levels = c('D7', 'D6', 'D4', 'D28', 'D26', 'D19', 'MD10', 'P8', 'C10', 'C3')),
    ztest = ifelse(ztest < 0.05, '(p < 0.05)', '(ns)'),
    resvar = factor(resvar, levels = reslabs$shrt, labels = reslabs$expr), 
    trndvar = factor(trndvar, levels = c('res', 'norm'), labels = c('Obs.', 'Flow-norm.'))
  ) %>% 
  unite('trnd', trndvar, ztest, sep = ' ')

@@

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for aggregations by seasons from (a) 1976-1995 and (b) 1996-2013.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Station locations have been jittered to reduce overlap. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 5.3, cache = T, out.width = '0.7\\textwidth'>>=

@@

# combine all
trnds_seasyr <- rbind(trnds_nrm, trnds_fit, trnds_obs) %>% 
  select(Site_Code, resvar, cat, perchg, ann, trndvar) %>% 
  spread(trndvar, perchg) %>% 
  mutate(Site_Code = factor(Site_Code, levels = c('D7', 'D6', 'D4', 'D28', 'D26', 'D19', 'MD10', 'P8', 'C10', 'C3')))

@@

# need to shift locations of some stations
statmeta[statmeta$Site_Code %in% 'D26', 'Latitude'] <- 38.14
statmeta[statmeta$Site_Code %in% 'D26', 'Longitude'] <--121.55
statmeta[statmeta$Site_Code %in% 'D7', 'Latitude'] <- 38.13
statmeta[statmeta$Site_Code %in% 'D7', 'Longitude'] <--122.015
statmeta[statmeta$Site_Code %in% 'C10', 'Latitude'] <-37.68
statmeta[statmeta$Site_Code %in% 'C10', 'Longitude'] <--121.28
statmeta[statmeta$Site_Code %in% 'D28', 'Latitude'] <-37.94
statmeta[statmeta$Site_Code %in% 'MD10', 'Latitude'] <-38.1
statmeta[statmeta$Site_Code %in% 'P8', 'Latitude'] <-37.93
statmeta[statmeta$Site_Code %in% 'D19', 'Latitude'] <-38.06

@@

# bounds on limits
yrng <- max(abs(range(trnds_seasyr$norm)))
yrng <- c(-1 * yrng, yrng)

@@

      nms <- names(bars)[i]
      dat <- statmeta[statmeta$Site_Code %in% nms,]
      pall <- subview(pall, bars[[i]], x = dat$Longitude, y = dat$Latitude, subwid, 0.21) # change bar heights here

@@

# setup legend
rngs <- seq(yrng[1], yrng[2], length = length(c(colneg, colpos)))
leg <- ggplot(data.frame(x = seq_along(rngs), y = seq_along(rngs), z = rngs), aes(x, y, colour = z)) + 
  geom_point() + 
  scale_colour_gradientn(colours = c(rev(colneg), colpos)) + 
  guides(colour = guide_colourbar(title = expression('% chg yr' ^-1), barheight = 0.5, barwidth = 8)) + 
  theme(legend.position = 'top')
leg <- g_legend(leg)

@@

# dynaplots
p5 <- dynaplot(nomod, logspace = F, col_vec = dynacol_vec, month = c(1, 4, 7, 10), floscl = F, size = 0.5) + 
  scale_x_continuous(flolab) + 
  scale_y_continuous(ylab1) +
  ptheme %+replace% 
  theme(
    legend.position = 'top', 
    legend.title = element_blank()
  ) +
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10))
pleg2 <- g_legend(p5)
p5 <- p5 + ptheme %+replace% theme(legend.position = 'none')
p6 <- dynaplot(nhmod, logspace = F, col_vec = dynacol_vec, month = c(1, 4, 7, 10), floscl = F, size = 0.5) + 
  scale_x_continuous(flolab) +
  scale_y_continuous(ylab2) +
  ptheme %+replace% 
  theme(
    legend.position = 'none', 
    legend.title = element_blank()
  )

@@

<<clmchl, fig.height = 9, fig.width = 8, out.width = '0.9\\textwidth', cache = T, fig.cap = 'Trends in clam abundance and \\ac{chla} concentration from 1976 to 2013 at station D7 in Suisun Bay.  Invasion by \\textit{Potamocorbula amurensis} clams in the late 1980s and displacement of \\textit{Corbicula fluminea} was shown by changes in clam density (a, annual means),with biomass linked to salinity (b).  A coincident decrease in \\ac{chla} concentration was also observed by changes in annual (c) and seasonal trends(d).  A significant ($p < 0.001$) relationship between clam biomass and \\ac{chla} concentration is shown in subfigure (e).  Flow relationships with \\ac{chla} concentration have also changed over time (f, observations from June).'>>=
# clam and model data at D7
data(clams)
data(h3dat)
data(flow_dat)
chld7 <- filter(h3dat, Site_Code %in% 'D7' & resvar %in% 'chl')$mod[[1]]

@@

# colors 
cols <- RColorBrewer::brewer.pal(10, 'Set2')[1:2]

# theme
pmar <- grid::unit(c(0, 6, 0, 6), 'pt')
thjust <- 0; tvjust <- -10
mythm <- theme_minimal() + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    legend.position = 'top', 
    legend.title = element_blank(), 
    legend.box = 'horizontal',    
    plot.title = element_text(hjust = thjust, vjust = tvjust),
    plot.margin = pmar
    )

##
# annual clam time series
p1 <- ggplot(clmdat, aes(x = yr, y = clams_smp, fill = species)) + 
  geom_bar(stat = 'identity') + 
  mythm +
  scale_fill_brewer(
    palette = 'Set2',
    labels = parse(text = c('italic(Corbicula)', 'italic(Potamocorbula)'))
    ) + 
  scale_x_date('Year', limits = xlims) +
  scale_y_continuous('Clams per sample') +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  labs(title = '(a)')

@@

# scatter plot of concentration x biomass with model 
p2 <- ggplot(clmflo, aes(x = q, y = biomass)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  mythm + 
  scale_colour_manual(label = 'log(1 + biomass) = 5.2 - 0.6log(1 + flow)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0,15)) +
  labs(title = '(b)')

# prdnrmplot
ylab <- expression(paste("ln-chlorophyll (", italic(mu), "g ", L^-1, ")"))

p3 <- prdnrmplot(chld7, predicted = F, col_vec = cols, min_mo = 10) + 
  mythm + 
  scale_x_date('Year') +
  scale_y_continuous(ylab) +
  labs(title = '(c)')

# seasyrplot
p4 <- seasyrplot(chld7, predicted = F, col_vec = cols, min_mo = 10) + 
  theme(
    axis.line.x = element_line(), 
    axis.line.y = element_line(), 
    axis.ticks.x = element_line(),
    axis.ticks.y = element_line(),
    axis.ticks.length = unit(.1, "cm"), 
    legend.title = element_blank(), 
    panel.border = element_blank(), 
    axis.title.x = element_text(),
    plot.title = element_text(hjust = thjust, vjust = tvjust),
    plot.margin = pmar
    ) + 
  scale_x_date('Day of year', date_labels = "%m/%d") +
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10)) +
  scale_y_continuous(ylab) + 
  labs(title = '(d)')

##
# chlorophyll by month
chldat <- filter(h3dat, resvar %in% 'chl' & Site_Code %in% 'D7') %>% 
  .$mod %>% 
  .[[1]] %>% 
  select(year, month, res, bt_norm) %>% 
  mutate(res = exp(res)) %>% 
  rename(yr = year, mo = month, bt_res = res) %>% 
  group_by(yr, mo) %>% 
  dplyr::summarize(res = mean(bt_res, na.rm = TRUE), bt_norm = mean(bt_norm, na.rm = TRUE)) %>% 
  ungroup

@@

xlab <- expression(paste(italic(Potamocorbula), " biomass (g ", m^-2, ")"))

# scatter plot of concentration x biomass with model 
p5 <- ggplot(toplo, aes(x = biomass, y = bt_norm)) + 
  geom_point(size = 3, alpha = 0.8, colour = 'darkgrey') + 
  geom_line(data = data.frame(x = xvals, y = yvals), aes(x, y, colour = 'black'), lwd = 1) +
  mythm + 
  scale_colour_manual(label = 'log(1 + chl) = 1.35 - 0.1log(1 + biomass)', values = 'black') +
  scale_x_continuous(xlab) + 
  scale_y_continuous(ylab, limits = c(0.8, 4)) +
  labs(title = '(e)')

##
# dynaplot for chl, june
p6 <- dynaplot(chld7, mo = 6, col_vec = cols, floscl = F, size = 0.5) +
  mythm %+replace%
  theme(
    strip.background = element_blank(),
    strip.text.x = element_blank()
    ) + 
  scale_x_continuous('Salinity (psu)') + 
  scale_y_continuous(ylab, limits = c(0.4, 3.3)) +
  guides(colour = guide_colourbar(barheight = 0.5, barwidth = 10)) + 
  labs(title = '(f)')

##
# dimension formatting

# align widths of plots in first column
pA <- ggplot_gtable(ggplot_build(p1))
pB <- ggplot_gtable(ggplot_build(p3))
pC <- ggplot_gtable(ggplot_build(p5))
maxWidth = unit.pmax(pA$widths[2:3], pB$widths[2:3], pC$widths[c(2:3)])

# Set the widths
pA$widths[2:3] <- maxWidth
pB$widths[2:3] <- maxWidth
pC$widths[2:3] <- maxWidth

# align widths of plots in second column
pD <- ggplot_gtable(ggplot_build(p2))
pE <- ggplot_gtable(ggplot_build(p4))
pF <- ggplot_gtable(ggplot_build(p6))
maxWidth = unit.pmax(pD$widths[2:3], pE$widths[2:3], pF$widths[2:3])

# Set the widths
pD$widths[2:3] <- maxWidth
pE$widths[2:3] <- maxWidth
pF$widths[2:3] <- maxWidth

# combine all
grid.arrange(
  arrangeGrob(pA, pB, pC, ncol = 1, heights = c(1, 1, 1)), 
  arrangeGrob(pD, pE, pF, ncol = 1, heights = c(1, 1, 1)), 
  ncol = 2, widths = c(1, 1)
  )
@

@@

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Annual'),
  n.cgroup = c(2),
  rowname = stats,
  label = 'tab:trndsann',
  insert.bottom = foot.val
  )

@@

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Seasonal, 1976-1995'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsbef',
  insert.bottom = foot.val
  )

@@

# table
latex( 
  tab[, -c(1, 2)],
  file = '',
  rowlabel = 'Analyte/Station',
  caption = cap.val,
  caption.loc = 'top',
  rgroup = vars,
  n.rgroup = rep(nrow(tab)/3, 3),
  cgroup = c('Seasonal, 1996-2013'),
  n.cgroup = c(4),
  rowname = stats,
  label = 'tab:trndsaft',
  insert.bottom = foot.val
  )

@@
